Platform organizations increasingly deploy optimization algorithms under the banner of ``digital innovation'' or ``AI-driven transformation,'' when in practice they primarily optimize existing processes and metrics rather than fundamentally reimagining organizational possibilities \parencite{smith2023}. Recent implementations across major platforms exemplify this mischaracterization through multiple initiatives marketed as innovative solutions. Ride-sharing platforms promote their dispatch algorithms as ``innovative mobility solutions,'' yet these systems primarily optimize traditional transportation metrics---reducing wait times by 31\% and improving vehicle utilization by 26\% through conventional supply-demand matching \parencite{chen2022}. Similarly, content delivery platforms market their recommendation engines as ``revolutionary content discovery innovations,'' while fundamentally optimizing traditional engagement metrics, achieving 28\% higher view completion rates through refined content sequencing \parencite{johnson2024}.

The tension between innovation rhetoric and optimization practices creates profound challenges for platform stakeholders. Platform workers, attracted by narratives of ``innovative flexible work,'' discover their autonomy strictly bounded by optimization algorithms---ride-share drivers report 38\% less control over route selection, while content moderators face 44\% more rigid decision protocols \parencite{kellogg2020}. These findings reveal a consistent pattern: platform stakeholders enter optimization initiatives expecting innovation's expansive possibilities but encounter increasingly constrained experiences defined by algorithmic efficiency metrics.

These examples raise critical questions about the implementation of optimization systems in platform contexts. How do platforms effectively measure and balance optimization metrics against stakeholder impacts? What methods can validate both quantitative improvements and qualitative outcomes? How do optimization systems reshape professional roles and relationships within platform networks? This research addresses these questions through a systematic investigation of optimization system implementations across major platform contexts. 